---
title: "Data visualization in R: ggplot2"
author: "David Lähnemann"
date: "August 31, 2016"
output:
  ioslides_presentation:
    widescreen: true
    fig_width: 10
    fig_height: 4
---

# ggplot2: easy and beautiful

Hadley Wickham's (almost completely) hassle-free implementation of:

* ["The Grammar Of Graphics"](https://www.cs.uic.edu/~wilkinson/TheGrammarOfGraphics/GOG.html) by *Leland Wilkinson*.

## ggplot2 resources {.bigger}

### R Cookbook on Graphs:

* Getting started with all the basics, nice and crispy!
* http://www.cookbook-r.com/Graphs/

### Official docs:

* Extensive documentation, with lots of examples.
* http://docs.ggplot2.org/current/

## Setup: ggplot2 and example data

```{r, message = FALSE, quiet=TRUE}
install.packages("ggplot2")
library("ggplot2")
head(diamonds)
```

# Building plots


## Building blocks {.bigger}

### **geom:** a shape for data representation

### **stat:** a (statistical) data transformation

### **aes:** mapping aesthetics to data levels

### **coord:** coordinate systems

### **+** whatever customizations you might want


## qplot: **q**uick plots, smart defaults

```{r}
p <- qplot(diamonds$cut)
p + geom_bar()
```

## ggplot: more customization

```{r eval = FALSE}
p <- ggplot(data=diamonds)
p + geom_point()
```

Fails with:
```{r eval = FALSE}
Error: geom_point requires the following missing aesthetics: x, y
```

## Specifiy an aesthetics mapping
```{r}
p <- ggplot(data=diamonds, aes(x=carat,y=price))
p + geom_point()
```

## Leverage aesthetics mappings

```{r}
p <- ggplot(data=diamonds, aes(x=carat,y=price))
p + geom_point(aes(colour=color, shape=cut))
```

## Liven up that bar chart

```{r}
p <- ggplot(data=diamonds, aes(x=cut))
p + geom_bar(aes(fill=color))
```

## A boxplot: good-looking standards
```{r}
p <- ggplot(data=diamonds, aes(x=cut,y=depth))
p + geom_boxplot()
```

## See it all: easy layering
```{r}
p <- ggplot(data=diamonds, aes(x=cut,y=depth))
p + geom_jitter() + geom_boxplot()
```

## Cleaning up: easy tweaking

```{r}
p <- ggplot(data=diamonds, aes(x=cut,y=depth))
p + geom_jitter(alpha=0.3) + geom_boxplot(alpha=0.7)
```

## Turn it all around: the coordinate system

```{r}
p <- ggplot(data=diamonds, aes(x=cut,y=depth))
p + geom_jitter(alpha=0.3) + geom_boxplot(alpha=0.7) + coord_flip()
```

# ...geht fast von allein.

> Just do it...

# P.S.: Rmarkdown + ioslides

### Rmarkdown: http://rmarkdown.rstudio.com/
* https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf

### Rmarkdown ioslides:
* http://rmarkdown.rstudio.com/ioslides_presentation_format.html

## HowTo produce these slides

```{r eval = FALSE}
git clone https://github.com/dlaehnemann/ggplot2_Intro
cd ggplot2_Intro
# Edit ggplot2_Intro.Rmd
R
> install.packages("rmarkdown")
> library("rmarkdown")
> install.packages("knitr")
> library("knitr")
> render("ggplot2_Intro.Rmd")
q()
```
 Open the provided or your produced ggplot2\_Intro.html with any web browser.